### 2025-08-07 actuarial.qmd
20250807 - tweaks

Another potential reason clinicians may not use actuarial judgment is time.
This can be a legitimate challenge to using actuarial models in some contexts.
Actuarial methods can take time to apply—they require assessing the relevant criteria, putting the assessment scores into an actuarial formula, and interpreting the output score.
In some clinical situations, such as risk assessments in terms of whether someone is imminently suicidal or homicidal and how to proceed, clinicians must make immediate in-session decisions [@Jacinto2018].
Thus, it can be challenging or impractical in some situations to conduct all of the relevant assessments needed to apply actuarial approaches.
Moreover, in some workplace settings, such as health care settings, clinicians may be pushed beyond their limits in terms of productivity expectations.
In these situations, one possible approach is to exercise one's best possible judgment in the moment based on the available evidence [i.e., intuitive decision-making processes; @Jacinto2018], but to check it against an actuarial model after the fact.
This is like comparing your decisions in a chess match, after the game ends, to how a computer would have moved in each of the situations.
It is valuable to receive feedback on the accuracy of one's judgments and, when time permits, to evaluate one's intuitive decision-making against analytical decision-making.
However, there are structured approaches to assessing suicide risk that are quick to administer [e.g., @Posner2011].

@@

There are various challenges of data-driven approaches.
First, they are sometimes not interpretable or consistent with theory.
Second, they tend to [overfit](#sec-overfitting) the data.
[Overfitting](#sec-overfitting) is described in @sec-overfitting.
Third, as a result of [overfitting](#sec-overfitting) the data, they tend to show [shrinkage](#sec-shrinkage).

>>>>

Another potential reason clinicians may not use actuarial judgment is time.
This can be a legitimate challenge to using actuarial models in some contexts.
Actuarial methods can take time to apply—they require assessing the relevant criteria, putting the assessment scores into an actuarial formula, and interpreting the output score.
In some clinical situations, such as risk assessments in terms of whether someone is imminently suicidal or homicidal and how to proceed, clinicians must make immediate in-session decisions [@Jacinto2018].
Thus, it can be challenging or impractical in some situations to conduct all of the relevant assessments needed to apply actuarial approaches.
Moreover, in some workplace settings, such as health care settings, clinicians may be pushed beyond their limits in terms of productivity expectations.
In these situations, one possible approach is to exercise one's best possible judgment in the moment based on the available evidence [i.e., intuitive decision-making processes; @Jacinto2018], and to check it against an actuarial model after the fact.
This is like comparing your decisions in a chess match, after the game ends, to how a computer would have moved in each of the situations.
It is valuable to receive feedback on the accuracy of one's judgments and, when time permits, to evaluate one's intuitive decision-making against analytical decision-making.
However, there are structured approaches to assessing suicide risk that are quick to administer [e.g., @Posner2011].

@@

There are various challenges of data-driven approaches (e.g., machine learning).
First, they are sometimes not interpretable or consistent with theory.
Sometimes that may because are theories are wrong; other times, it may mean we need to refine our data-driven approach.
Second, data-driven approaches tend to [overfit](#sec-overfitting) the data.
[Overfitting](#sec-overfitting) is described in @sec-overfitting.
Third, as a result of [overfitting](#sec-overfitting) the data, data-driven approaches tend to show [shrinkage](#sec-shrinkage).
